---
title: "Session 5: {{< meta session-title >}}"
date: 2025-01-22T15:00:00
format: revealjs
metadata-files:
  - _slides.yml
session-title: "Branches"
chapter: |
  ["Branches"](https://lennartwittkuhn.com/version-control-book/chapters/branches.html)
quiz-url: https://version-control-branches.formr.org/
---

# Last session

## Last session: First steps with Git & Git Essentials

After the last session, you should now be able to answer the following questions / do the following:

::: fragment
### First steps with Git

{{< include ../objectives/_objectives-firststeps.qmd >}}
:::

::: {.fragment}
### Git essentials

{{< include ../objectives/_objectives-essentials.qmd >}}
:::

## Last session: Cheatsheet

### First steps with Git

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
knitr::kable(table_cheatsheet(name = "basic"), format = "markdown", row.names = FALSE)
```

## Last session: Cheatsheet

### Git essentials

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
knitr::kable(table_cheatsheet(name = "essentials"), format = "markdown", row.names = FALSE)
```

## Last session: `{{< var course.project >}}` project

### First steps with Git

{{< include project/project-basics.qmd >}}

## Last session: `{{< var course.project >}}` project

### Git Essentials

{{< include project/project-essentials.qmd >}}

## Common questions after last session

::: fragment
#### {{< fa circle-question >}} *"How can I make changes in a text (`.txt`) file?"*
- Remember: **It's just a regular text file on your computer.**
- Option 1: Use a regular text editor (e.g., TextEdit on macOS, Notepad++ on Windows, or others).
- Option 2: Use `vim` directly from the command line (see chapter / session on "Setup").
- Option 3: Use command line commands like `echo "My text" >> {{< var course.project >}}.txt`.
:::

::: fragment
#### {{< fa circle-question >}} *"I ran `git commit` but nothing happened?"*
- Maybe you forgot to run `git add` before?
:::

::: fragment

#### {{< fa circle-exclamation >}} *"Remember to `cd` into your `{{< var course.project >}}` project folder!"*

- Every time you open the command line, it starts at a default location (likely your home directory).
- Check in which folder you are, using `pwd` ("print working directory").
- Use `cd` to move to your `{{< var course.project >}}` project folder before you start running the Git commands.

:::

{{< include _git-home-directory.qmd >}}

## Staging and Committing

::: columns
::: {.column width="50%"}
### Staging:

-   Command: `git add`
-   Purpose: Preparing and organizing files before they are recorded in the repository's history.
:::

::: {.column width="50%"}
### Committing:

-   Command: `git commit -m "commit message"`
-   Purpose: Saving the changes in the staged files to the repository's history, creating a snapshot.
:::
:::

::: columns
::: {.column width="50%"}
::: fragment
![Source: [git-scm.com](https://git-scm.com/book/en/v2)](../images/git-staging-diagram.png){width="100%"}
:::
:::
::: {.column width="50%"}
::: fragment
![Source: [Heidi Seibold](https://docs.google.com/presentation/d/1Yj59XAAJZQprO-kfDHzHMtjJDttMFAqkkslqcSIRZKg/edit#slide=id.p)](../images/git-analogy-letter.png){width="100%"}
:::
:::
:::

# This session: {{< meta session-title >}}

## This session: {{< meta session-title >}}

![Image from [Chapter "Git Branches"](https://book.the-turing-way.org/reproducible-research/vcs/vcs-git-branches.html) of the ["The Turing Way handbook to reproducible, ethical and collaborative data science"](https://book.the-turing-way.org/), used under a [Creative Commons Attribution 4.0 License](https://github.com/the-turing-way/the-turing-way/blob/main/LICENSE.md).](../images/turing-way-branches.png){#fig-branches}

## Learning objectives

{{< include ../objectives/_objectives-branches.qmd >}}

## Reading

{{< fa display >}} {{< meta chapter >}}

```{=html}
<iframe width="100%" height="80%" src="https://lennartwittkuhn.com/version-control-book/chapters/branches.html"></iframe>
```

::: {.notes}
**Illustration of branches in a Git repository.** Each circle represents a commit. Commits on the `Main` branch are shown in gray. The branch `Feature A` (blue circles) was created from the `Main` branch and later merged again. In-between, another branch `Feature A-1` was created from the `Feature A` branch (green circles) and merged with that branch again before the `Feature A` branch was eventually merged with the `Main` branch. The branch `Feature B` was created but abandoned and not merged with the `Main` branch. 
:::

## Tasks

{{< include tasks/tasks-template.qmd >}}

## `{{< var course.project >}}` project

{{< include project/project-branches.qmd >}}

## Cheatsheet

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
knitr::kable(table_cheatsheet(name = "branches"), format = "markdown", row.names = FALSE)
```

## Exercises

### Branches

{{< include ../exercises/_exercises-branches.qmd >}}

## :rocket: Bonus exercises

### Create and resolve a merge conflict

{{< include ../exercises/_exercises-branches-merge-conflict.qmd >}}

## Solutions: Branches

```{zsh, filename="Code"}
#| eval: false
#| file: exercises/code-exercises-branches-recipes.sh
```
{{< include ../exercises/_solutions-branches.qmd >}}

## Solutions: Merge Conflict

```{zsh, filename="Code"}
#| eval: false
#| file: exercises/code-exercises-branches-merge-conflict.sh
```
{{< include ../exercises/_solutions-branches-merge-conflict.qmd >}}
